---
title: "Three plots from amino acid group additivity paper (2006)"
output: 
  rmarkdown::html_vignette:
    mathjax: null
vignette: >
  %\VignetteIndexEntry{Three plots from amino acid group additivity paper (2006)}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r HTML, include=FALSE}
## some frequently used HTML expressions
pKa <- "p<i>K</i><sub>a</sub>"
```

This vignette from the R package [JMDplots](https://github.com/jedick/JMDplots) contains one plot from the paper by <a href="https://doi.org/10.5194/bg-3-311-2006" title="Temperature, pressure, and electrochemical constraints on protein speciation: Group additivity calculation of the standard molal thermodynamic properties of ionized unfolded proteins">Dick et al. (2006)</a>.

```{r JMDplots, message = FALSE, results = "hide"}
library(JMDplots)
```

## `r pKa` of sidechain and amino acid backbone groups (Figure 4)
```{r aaaq4, message = FALSE, echo = 1, fig.align='center', out.width='80%', out.height='450px'}
aaaq4(TRUE)
knitr::include_graphics("aaaq4.pdf")
```

## Net charge of proteins as a function of pH and *T* (Figure 10)
```{r aaaq10, message = FALSE, echo = 1, fig.align='center', out.width='90%', out.height='550px'}
aaaq10(TRUE)
knitr::include_graphics("aaaq10.pdf")
```

## Eh-pH diagram for extracellular α-amylases (Figure 13)
This plot, showing the relative stabilities of extracellular α-amylases from two organisms on an Eh-pH diagram, is based on Fig. 13 of Dick et al. (2006).
Points are added for Eh-pH values from soils ([Baas Becking et al., 1960](https://www.jstor.org/stable/30059218)) and hot springs in Yellowstone ([Spear et al., 2005](https://doi.org/10.1073.pnas.0409574102)) and Iceland ([Stefánsson and Arnórsson, 2002](https://doi.org/10.1016/S0009-2541(02)00119-5)).
Note that the symbols for Yellowstone and Iceland were incorrectly swapped in the figure caption of Dick et al. (2006).

```{r aaaq13, message = FALSE, echo = 1, fig.align='center', out.width='80%', out.height='480px'}
aaaq13(TRUE)
knitr::include_graphics("aaaq13.pdf")
```
