---
title: "Amino acid group additivity (2006)"
output: 
  rmarkdown::html_vignette:
    mathjax: null
vignette: >
  %\VignetteIndexEntry{Amino acid group additivity (2006)}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
library(knitr)
## use pngquant to reduce size of PNG images
knit_hooks$set(pngquant = hook_pngquant)
pngquant <- "--speed=1 --quality=0-25"
# in case pngquant isn't available
if (!nzchar(Sys.which("pngquant"))) pngquant <- NULL 
```

```{r HTML, include=FALSE}
## some frequently used HTML expressions
pKa <- "p<i>K</i><sub>a</sub>"
```

This vignette runs the code to make selected plots from this paper:

> Dick JM, LaRowe DE, Helgeson HC. 2006. Temperature, pressure, and electrochemical constraints on protein speciation: Group additivity calculation of the standard molal thermodynamic properties of ionized unfolded proteins. *Biogeosciences* **3**(3): 311–336. doi: [10.5194/bg-3-311-2006](https://doi.org/10.5194/bg-3-311-2006)

These plots have been modified from the original figures to add a neutral pH line to Figure 4 and a legend and colored labels to Figure 13.

This vignette was compiled on `r Sys.Date()` with **[JMDplots](https://github.com/jedick/JMDplots)** `r packageDescription("JMDplots")$Version` **[CHNOSZ](http://chnosz.net)** `r packageDescription("CHNOSZ")$Version`.

```{r JMDplots, message = FALSE, results = "hide"}
library(JMDplots)
```

## `r pKa` of sidechain and amino acid backbone groups (Figure 4)
```{r aaaq4, message = FALSE, fig.align='center', out.width='80%', fig.width = 8, fig.height = 6, pngquant = pngquant}
aaaq4()
```

## Net charge of proteins as a function of pH and *T* (Figure 10)
```{r aaaq10, message = FALSE, fig.align='center', out.width='90%', fig.width = 6, fig.height = 5, pngquant = pngquant}
aaaq10()
```

## Eh-pH diagram for extracellular α-amylases (Figure 13)
This plot, showing the relative stabilities of extracellular α-amylases from two organisms on an Eh-pH diagram, is based on Fig. 13 of Dick et al. (2006).
Points are added for Eh-pH values from soils ([Baas Becking et al., 1960](https://www.jstor.org/stable/30059218)) and hot springs in Yellowstone ([Spear et al., 2005](https://doi.org/10.1073/pnas.0409574102)) and Iceland ([Stefánsson and Arnórsson, 2002](https://doi.org/10.1016/S0009-2541(02)00119-5)).
Note that the symbols for Yellowstone and Iceland were incorrectly swapped in the figure caption of Dick et al. (2006).

```{r aaaq13, message = FALSE, fig.align='center', out.width='80%', fig.width = 5, fig.height = 4, pngquant = pngquant}
aaaq13()
```
