---
title: "Seven plots from oxidation state of carbon paper (2014)"
output: 
  rmarkdown::html_vignette:
    mathjax: null
vignette: >
  %\VignetteIndexEntry{Seven plots from oxidation state of carbon paper (2014)}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
library(knitr)
## use pngquant to reduce size of PNG images
knit_hooks$set(pngquant = hook_pngquant)
pngquant <- "--speed=1 --quality=0-25"
# in case pngquant isn't available
if (!nzchar(Sys.which("pngquant"))) pngquant <- NULL 
```

```{r HTML, include=FALSE}
## some frequently used HTML expressions
ZC <- "<i>Z</i><sub>C</sub>"
Topt <- "<i>T</i><sub>opt</sub>"
```

This vignette from the R package [JMDplots](https://github.com/jedick/JMDplots) contains plots from the paper by <a href="https://doi.org/10.1098/rsif.2013.1095" title="Average oxidation state of carbon in proteins">Dick (2014)</a>.

```{r JMDplots, message = FALSE, results = "hide"}
library(JMDplots)
```

## `r ZC` of amino acids vs `r ZC` of RNA codons and hydropathy index of amino acids (Figure 1)
```{r aoscp1, message = FALSE, fig.align='center', out.width='100%', fig.width = 10, fig.height = 5, pngquant = pngquant}
aoscp1()
```

## Histograms of `r ZC` of all human proteins and human membrane proteins (Figure 2)
```{r aoscp2, message = FALSE, fig.align='center', out.width='100%', fig.width = 6.6, fig.height = 4.4, pngquant = pngquant}
aoscp2()
```

## Drawing of yeast cell color-coded with median `r ZC` of proteins in different locations (Figure 3)
```{r aoscp3, message = FALSE, fig.align='center', out.width='100%', fig.width = 12, fig.height = 8, pngquant = pngquant}
# first make the base plot, then add labels
aoscp3(TRUE)
aoscp3()
```

## `r ZC` and Eh ranges in yeast and ER-cytoplasm electron-transfer scheme (Figure 4)
```{r aoscp4, message = FALSE, fig.align='center', out.width='100%', fig.width = 8, fig.height = 4, pngquant = pngquant}
aoscp4()
```

## Average oxidation state of carbon in proteins from different organisms (Figure 5)
```{r aoscp5, message = FALSE, fig.align='center', out.width='100%', fig.width = 10, fig.height = 6, pngquant = pngquant}
aoscp5()
```

## `r ZC` and `r Topt` of different rubiscos and thermodynamic comparison (Figure 6)
```{r aoscp6, message = FALSE, fig.align='center', out.width='100%', fig.width = 8, fig.height = 5, pngquant = pngquant}
aoscp6()
```

## `r ZC` of ferredoxin, thioredoxin, and glutaredoxin vs midpoint reduction potential (unpublished)
```{r aoscp99, message = FALSE, fig.align='center', out.width='75%', fig.width = 6, fig.height = 5, pngquant = pngquant}
aoscp99()
```

