---
title: "Plots from oxidation state of carbon paper (2014)"
output: 
  rmarkdown::html_vignette:
    mathjax: null
vignette: >
  %\VignetteIndexEntry{Plots from oxidation state of carbon paper (2014)}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r HTML, include=FALSE}
## some frequently used HTML expressions
ZC <- "<i>Z</i><sub>C</sub>"
```

This vignette from the R package JMDplots contains plots from the paper by [Dick (2014)](https://doi.org/10.1098/rsif.2013.1095).

```{r JMDplots, message = FALSE, results = "hide"}
library(JMDplots)
```

## `r ZC` of amino acids vs `r ZC` of RNA codons and hydropathy index of amino acids (Figure 1)
```{r aoscp1, message = FALSE, echo = 1, fig.align='center', out.width='100%', out.height='350px'}
aoscp1(TRUE)
knitr::include_graphics("aoscp1.pdf")
```

## Histograms of `r ZC` of all human proteins and human membrane proteins (Figure 2)
```{r aoscp2, message = FALSE, echo = 1, fig.align='center', out.width='100%', out.height='450px'}
aoscp2(TRUE)
knitr::include_graphics("aoscp2.pdf")
```

## Drawing of yeast cell color-coded with median `r ZC` of proteins in different locations (Figure 3)
```{r aoscp3, message = FALSE, echo = 1:3, fig.align='center', out.width='100%', out.height='450px'}
# first make a PNG with colored areas, then PDF with labels
aoscp3(TRUE)
aoscp3()
knitr::include_graphics("aoscp3.pdf")
```

## `r ZC` and Eh ranges in yeast and ER-cytoplasm electron-transfer scheme (Figure 4)
```{r aoscp4, message = FALSE, echo = 1, fig.align='center', out.width='100%', out.height='400px'}
aoscp4(TRUE)
knitr::include_graphics("aoscp4.pdf")
```

## Average oxidation state of carbon in proteins from different organisms (Figure 5)
```{r aoscp5, message = FALSE, echo = 1, fig.align='center', out.width='100%', out.height='420px'}
aoscp5(TRUE)
knitr::include_graphics("aoscp5.pdf")
```

