---
title: "CHNOSZ supplementary plots (2019)"
output: 
  rmarkdown::html_vignette:
    mathjax: null
vignette: >
  %\VignetteIndexEntry{CHNOSZ supplementary plots (2019)}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
library(knitr)
## use pngquant to reduce size of PNG images
knit_hooks$set(pngquant = hook_pngquant)
pngquant <- "--speed=1 --quality=0-25"
# in case pngquant isn't available
if (!nzchar(Sys.which("pngquant"))) pngquant <- NULL 
```

This vignette runs the code to make most of the supplementary figures from the following paper:

> Dick JM. 2019. CHNOSZ: Thermodynamic calculations and diagrams for geochemistry. *Frontiers in Earth Science* **7**: 180. doi: [10.3389/feart.2019.00180](https://doi.org/10.3389/feart.2019.00180)

This vignette was compiled on `r Sys.Date()` with **[JMDplots](https://github.com/jedick/JMDplots)** `r packageDescription("JMDplots")$Version` and **[CHNOSZ](http://chnosz.net)** `r packageDescription("CHNOSZ")$Version`.

```{r JMDplots, message = FALSE, results = "hide"}
library(JMDplots)
```

## Comparison of logK and maximum affinity methods (Figure S1)
```{r chnosz10S1, message = FALSE, fig.align='center', out.width='75%', fig.width = 6.5, fig.height = 5, pngquant = pngquant}
chnosz10S1()
```

## Eh-pH diagram like Fig. 5A of [Caporuscio et al. (2017)](https://doi.org/10.1016/j.jnucmat.2016.12.036) (Figure S2)
```{r chnosz10S2, message = FALSE, fig.align='center', out.width='75%', fig.width = 4, fig.height = 4, pngquant = pngquant}
chnosz10S2()
```

## Gibbs energy of transformation for an assemblage of n-alkanes (Figure S3)
```{r chnosz10S3, message = FALSE, fig.align='center', out.width='100%', fig.width = 7, fig.height = 4, pngquant = pngquant}
chnosz10S3()
```

## `findit()` calculations for sulfur species (Figure S4)
*This feature and figure were removed in CHNOSZ 2.0.0 and JMDplots 1.2.17, respectively.
Please refer to earlier versions of the packages.*

## Debye-Hückel extended term parameter extrapolated from plots of [Manning et al. (2013)](https://doi.org/10.2138/rmg.2013.75.5) (Figure S5)
```{r chnosz10S5, message = FALSE, fig.align='center', out.width='75%', fig.width = 6, fig.height = 6, pngquant = pngquant}
chnosz10S5()
```

## Figure 6 modified to exclude DEW data for acetate (Figure S6A)
```{r chnosz10S6A, message = FALSE, fig.align='center', out.width='75%', fig.width = 5, fig.height = 5, pngquant = pngquant}
chnosz10S6A()
```

## Figure S6A modified to use default bgamma equation (Figure S6B)
```{r chnosz10S6B, message = FALSE, fig.align='center', out.width='75%', fig.width = 5, fig.height = 5, pngquant = pngquant}
chnosz10S6B()
```

## log<i>K</i> of NaCl dissociation (Figure S7)
```{r chnosz10S7, message = FALSE, warning = FALSE, fig.align='center', out.width='75%', fig.width = 7, fig.height = 7, pngquant = pngquant}
chnosz10S7()
```

## Calcite solubility: comparison with [Manning et al. (2013)](https://doi.org/10.2138/rmg.2013.75.5) (Figure S8)
```{r chnosz10S8, message = FALSE, fig.align='center', out.width='75%', fig.width = 7, fig.height = 7, pngquant = pngquant}
chnosz10S8()
```

## Compare gold solubility in HCh and CHNOSZ: hematite-magnetite buffer (Figure S9)
```{r chnosz10S9, message = FALSE, fig.align='center', out.width='100%', fig.width = 9, fig.height = 7, pngquant = pngquant}
chnosz10S9()
```

## Compare gold solubility in HCh and CHNOSZ: pyrite-pyrrhotite-magnetite buffer (Figure S10)
```{r chnosz10S10, message = FALSE, fig.align='center', out.width='100%', fig.width = 9, fig.height = 7, pngquant = pngquant}
chnosz10S10()
```

