---
title: "Supplementary Plots from CHNOSZ paper"
output: 
  rmarkdown::html_vignette:
    mathjax: null
vignette: >
  %\VignetteIndexEntry{Supplementary Plots from CHNOSZ paper}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

This vignette from the R package JMDplots contains supplementary plots from the paper by <a href="https://doi.org/10.3389/feart.2019.00180" title="CHNOSZ: Thermodynamic calculations and diagrams for geochemistry">Dick (2019)</a>.

```{r JMDplots, message = FALSE, results = "hide"}
library(JMDplots)
```

## Comparison of logK and maximum affinity methods (Figure S1)
```{r chnsztenS1, message = FALSE, echo = 1, fig.align='center', out.width='50%', out.height='300px'}
chnsztenS1()
knitr::include_graphics("chnsztenS1.pdf")
```

## Eh-pH diagram like Fig. 5A of [Caporuscio et al. (2017)](https://doi.org/10.1016/j.jnucmat.2016.12.036) (Figure S2)
```{r chnsztenS2, message = FALSE, echo = 1, fig.align='center', out.width='50%', out.height='350px'}
chnsztenS2()
knitr::include_graphics("chnsztenS2.pdf")
```

## Gibbs energy of transformation for an assemblage of n-alkanes (Figure S3)
```{r chnsztenS3, message = FALSE, echo = 1, fig.align='center', out.width='100%', out.height='300px'}
chnsztenS3()
knitr::include_graphics("chnsztenS3.pdf")
```

## `findit()` calculations for sulfur species (Figure S4)
```{r chnsztenS4, message = FALSE, echo = 1, fig.align='center', out.width='100%', out.height='350px'}
chnsztenS4()
knitr::include_graphics("chnsztenS4.pdf")
```

## Debye-Hückel extended term parameter extrapolated from plots of [Manning et al. (2013)](https://doi.org/10.2138/rmg.2013.75.5) (Figure S5)
```{r chnsztenS5, message = FALSE, echo = 1, fig.align='center', out.width='50%', out.height='300px'}
chnsztenS5()
knitr::include_graphics("chnsztenS5.pdf")
```

## `chnszten6()` modified to exclude DEW data for acetate (Figure S6A)
## `chnsztenS6A()` modified to use default bgamma equation (Figure S6B)
```{r chnsztenS6, message = FALSE, echo = 1:2, fig.align='center', out.width='100%', out.height='300px'}
chnsztenS6A()
chnsztenS6B()
```

<!-- side-by-side images: https://stackoverflow.com/questions/39173714/r-markdown-can-i-insert-a-pdf-to-the-r-markdown-file-as-an-image -->
![Image1](chnsztenS6A.pdf){width=50%}![Image2](chnsztenS6B.pdf){width=50%}

## log<i>K</i> of NaCl dissociation (Figure S7)
```{r chnsztenS7_calc, message = FALSE, results = "hide", warning = FALSE}
chnsztenS7()
```
```{r chnsztenS7, message = FALSE, echo = FALSE, fig.align='center', out.width='50%', out.height='350px'}
knitr::include_graphics("chnsztenS7.pdf")
```

## Calcite solubility: comparison with [Manning et al. (2013)](https://doi.org/10.2138/rmg.2013.75.5) (Figure S8)
```{r chnsztenS8, message = FALSE, echo = 1, fig.align='center', out.width='50%', out.height='350px'}
chnsztenS8()
knitr::include_graphics("chnsztenS8.pdf")
```

## Compare gold solubility in HCh and CHNOSZ: hematite-magnetite buffer (Figure S9)
```{r chnsztenS9, message = FALSE, echo = 1, fig.align='center', out.width='100%', out.height='350px'}
chnsztenS9()
knitr::include_graphics("chnsztenS9.pdf")
```

## Compare gold solubility in HCh and CHNOSZ: pyrite-pyrrhotite-magnetite buffer (Figure S10)
```{r chnsztenS10, message = FALSE, echo = 1, fig.align='center', out.width='100%', out.height='350px'}
chnsztenS10()
knitr::include_graphics("chnsztenS10.pdf")
```

