---
title: "Ten supplementary plots from CHNOSZ paper (2019)"
output: 
  rmarkdown::html_vignette:
    mathjax: null
vignette: >
  %\VignetteIndexEntry{Ten supplementary plots from CHNOSZ paper (2019)}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
library(knitr)
## use pngquant to reduce size of PNG images
knit_hooks$set(pngquant = hook_pngquant)
pngquant <- "--speed=1 --quality=0-25"
# in case pngquant isn't available
if (!nzchar(Sys.which("pngquant"))) pngquant <- NULL 
```

This vignette from the R package [JMDplots](https://github.com/jedick/JMDplots) version `r packageDescription("JMDplots")$Version` contains supplementary plots from the paper by <a href="https://doi.org/10.3389/feart.2019.00180" title="CHNOSZ: Thermodynamic calculations and diagrams for geochemistry">Dick (2019)</a>.

```{r JMDplots, message = FALSE, results = "hide"}
library(JMDplots)
```

## Comparison of logK and maximum affinity methods (Figure S1)
```{r chnsztenS1, message = FALSE, fig.align='center', out.width='75%', fig.width = 6.5, fig.height = 5, pngquant = pngquant}
chnsztenS1()
```

## Eh-pH diagram like Fig. 5A of [Caporuscio et al. (2017)](https://doi.org/10.1016/j.jnucmat.2016.12.036) (Figure S2)
```{r chnsztenS2, message = FALSE, fig.align='center', out.width='75%', fig.width = 4, fig.height = 4, pngquant = pngquant}
chnsztenS2()
```

## Gibbs energy of transformation for an assemblage of n-alkanes (Figure S3)
```{r chnsztenS3, message = FALSE, fig.align='center', out.width='100%', fig.width = 7, fig.height = 4, pngquant = pngquant}
chnsztenS3()
```

## `findit()` calculations for sulfur species (Figure S4)
```{r chnsztenS4, message = FALSE, fig.align='center', out.width='100%', fig.width = 8, fig.height = 8, pngquant = pngquant}
chnsztenS4()
```

## Debye-Hückel extended term parameter extrapolated from plots of [Manning et al. (2013)](https://doi.org/10.2138/rmg.2013.75.5) (Figure S5)
```{r chnsztenS5, message = FALSE, fig.align='center', out.width='75%', fig.width = 6, fig.height = 6, pngquant = pngquant}
chnsztenS5()
```

## Figure 6 modified to exclude DEW data for acetate (Figure S6A)
```{r chnsztenS6A, message = FALSE, fig.align='center', out.width='75%', fig.width = 5, fig.height = 5, pngquant = pngquant}
chnsztenS6A()
```

## Figure S6A modified to use default bgamma equation (Figure S6B)
```{r chnsztenS6B, message = FALSE, fig.align='center', out.width='75%', fig.width = 5, fig.height = 5, pngquant = pngquant}
chnsztenS6B()
```

## log<i>K</i> of NaCl dissociation (Figure S7)
```{r chnsztenS7, message = FALSE, warning = FALSE, fig.align='center', out.width='75%', fig.width = 7, fig.height = 7, pngquant = pngquant}
chnsztenS7()
```

## Calcite solubility: comparison with [Manning et al. (2013)](https://doi.org/10.2138/rmg.2013.75.5) (Figure S8)
```{r chnsztenS8, message = FALSE, fig.align='center', out.width='75%', fig.width = 7, fig.height = 7, pngquant = pngquant}
chnsztenS8()
```

## Compare gold solubility in HCh and CHNOSZ: hematite-magnetite buffer (Figure S9)
```{r chnsztenS9, message = FALSE, fig.align='center', out.width='100%', fig.width = 9, fig.height = 7, pngquant = pngquant}
chnsztenS9()
```

## Compare gold solubility in HCh and CHNOSZ: pyrite-pyrrhotite-magnetite buffer (Figure S10)
```{r chnsztenS10, message = FALSE, fig.align='center', out.width='100%', fig.width = 9, fig.height = 7, pngquant = pngquant}
chnsztenS10()
```

