---
title: "CHNOSZ paper (2019)"
output: 
  rmarkdown::html_vignette:
    mathjax: null
vignette: >
  %\VignetteIndexEntry{CHNOSZ paper (2019)}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
library(knitr)
## use pngquant to reduce size of PNG images
knit_hooks$set(pngquant = hook_pngquant)
pngquant <- "--speed=1 --quality=0-25"
# in case pngquant isn't available
if (!nzchar(Sys.which("pngquant"))) pngquant <- NULL 
```

This vignette runs the code to make selected plots from this paper:

> Dick JM. 2019. CHNOSZ: Thermodynamic calculations and diagrams for geochemistry. *Frontiers in Earth Science* **7**: 180. doi: [10.3389/feart.2019.00180](https://doi.org/10.3389/feart.2019.00180)

This vignette was compiled on `r Sys.Date()` with **[JMDplots](https://github.com/jedick/JMDplots)** `r packageDescription("JMDplots")$Version` and **[CHNOSZ](http://chnosz.net)** `r packageDescription("CHNOSZ")$Version`.

```{r JMDplots, message = FALSE, results = "hide"}
library(JMDplots)
```

## Mosaic Eh-pH diagram for the Cu-S-Cl-O-H system (Figure 4)
```{r chnszten4, message = FALSE, fig.align='center', out.width='100%', fig.width = 8, fig.height = 3, pngquant = pngquant}
chnszten4()
```

## Corundum solubility diagram (Figure 5)
```{r chnszten5, message = FALSE, fig.align='center', out.width='50%', fig.width = 4, fig.height = 3, pngquant = pngquant}
chnszten5()
```

## <a href="http://www.dewcommunity.org/" title="Deep Earth Water Community">DEW</a> model (based on Fig. 3 of [Sverjensky et al., 2014](https://doi.org/10.1038/ngeo2291)) (Figure 6)
```{r chnszten6, message = FALSE, fig.align='center', out.width='50%', fig.width = 5, fig.height = 5, pngquant = pngquant}
chnszten6()
```

## Thermodynamic properties of reactions involving Al-bearing minerals (Figure 7)
```{r chnszten7, message = FALSE, fig.align='center', out.width='100%', fig.width = 7.2, fig.height = 7.2, pngquant = pngquant}
chnszten7()
```
