---
title: "gradox plots"
output: 
  rmarkdown::html_vignette:
    mathjax: null
vignette: >
  %\VignetteIndexEntry{gradox plots}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

```{r HTML, include=FALSE}
## some frequently used HTML expressions
# use lowercase here because these tend to be variable names in the examples
zc <- "<i>Z</i><sub>C</sub>"
```

This vignette from the R package JMDplots contains plots from the paper by [Dick et al. (2019)](https://doi.org/10.3389/fmicb.2019.00120).
For a closer reproduction of the figures in the paper, all plots here are made as PDF files (TRUE in arguments).

```{r JMDplots, message = FALSE, results = "hide"}
library(JMDplots)
```

## Some general characteristics of `r zc` of DNA, RNA, and proteins (Figure 1)
```{r gradox1, message = FALSE, echo = 1, fig.align='center', out.width='100%', out.height='550px'}
gradox1(TRUE)
knitr::include_graphics("gradox1.pdf")
```

## Selected plots of DNA, RNA, and protein `r zc` (Figure 2)
```{r gradox2, message = FALSE, echo = 1, fig.align='center', out.width='100%', out.height='400px'}
gradox2(TRUE)
knitr::include_graphics("gradox2.pdf")
```

## All plots of DNA and RNA `r zc` (Figure S1)
```{r gradoxS1, message = FALSE, echo = 1, fig.align='center', out.width='100%', out.height='200px'}
mout <- gradoxS1(TRUE)
knitr::include_graphics("gradoxS1.pdf")
```

## All plots of protein `r zc` (Figure S2)
```{r gradoxS2, message = FALSE, echo = 1, fig.align='center', out.width='100%', out.height='200px'}
pout <- gradoxS2(TRUE)
knitr::include_graphics("gradoxS2.pdf")
```

## `r zc` of proteins vs DNA (metagenomes and metatranscriptomes) (Figure 3)
```{r gradox3, message = FALSE, echo = 1, fig.align='center', out.width='100%', out.height='500px'}
gradox3(mout, pout, TRUE)
knitr::include_graphics("gradox3.pdf")
```

## Thermodynamic calculations of relative stabilities along redox gradients (Figure 4)
```{r gradox4, message = FALSE, echo = 1, fig.align='center', out.width='100%', out.height='500px'}
gradox4(mout, pout, TRUE)
knitr::include_graphics("gradox4.pdf")
```

## `r zc` of reads classified to selected abundant species (Figure 5)
```{r gradox5, message = FALSE, echo = 1, fig.align='center', out.width='100%', out.height='500px'}
gradox5(TRUE)
knitr::include_graphics("gradox5.pdf")
```

## `r zc` of reads classified to selected abundant species (including deeper samples) (Figure S3)
```{r gradoxS3, message = FALSE, echo = 1, fig.align='center', out.width='100%', out.height='500px'}
gradox5(TRUE, NULL)
knitr::include_graphics("gradoxS3.pdf")
```

