---
title: "Six plots from salinity gradients paper (2019)"
output: 
  rmarkdown::html_vignette:
    mathjax: null
vignette: >
  %\VignetteIndexEntry{Six plots from salinity gradients paper (2019)}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r options, include=FALSE}
options(width = 40)
```

```{r HTML, include=FALSE}
## some frequently used HTML expressions
ZC <- "<i>Z</i><sub>C</sub>"
nH2O <- "<i>n</i><sub>H<sub>2</sub>O</sub>"
```

This vignette from the R package [JMDplots](https://github.com/jedick/JMDplots) contains plots from the paper by <a href="gradH2O.html" title="Geochemistry of metagenomes: Distinct compositional signatures of environmental redox and salinity gradients">Dick et al. (2019)</a> (preprint).

```{r JMDplots, message = FALSE, results = "hide"}
library(JMDplots)
```

## Number of reactions for each amino acid in _E. coli_ metabolic reconstruction
Reaction equations used for this calculation are from <a href="https://doi.org/10.1038/msb4100155" title="A genome-scale metabolic reconstruction for Escherichia coli K-12 MG1655 that accounts for 1260 ORFs and thermodynamic information">Feist et al. (2007)</a>.
```{r gradH2O0, message = FALSE, echo = 1}
gradH2O0()
```

## Basis species comparison and relative water content (Figure 1)
```{r gradH2O1, message = FALSE, echo = 1, fig.align='center', out.width='100%', out.height='500px'}
gradH2O1(TRUE)
knitr::include_graphics("gradH2O1.pdf")
```

## `r ZC` for selected redox gradients (Figure 2)
```{r gradH2O2, message = FALSE, echo = 1, fig.align='center', out.width='100%', out.height='450px'}
gradH2O2(TRUE)
knitr::include_graphics("gradH2O2.pdf")
```

## `r nH2O`-`r ZC` scatterplots for redox gradients and the Baltic Sea (Figure 3)
```{r gradH2O3, message = FALSE, echo = 1, fig.align='center', out.width='100%', out.height='350px'}
gradH2O3(TRUE)
knitr::include_graphics("gradH2O3.pdf")
```

## `r nH2O` for Baltic Sea metagenome and metatranscriptome in different size fractions (Figure 4)
```{r gradH2O4, message = FALSE, echo = 1, fig.align='center', out.width='100%', out.height='350px'}
gradH2O4(TRUE)
knitr::include_graphics("gradH2O4.pdf")
```

## `r nH2O`-`r ZC` for freshwater, marine, and hypersaline environments (Figure 5)
```{r gradH2O5, message = FALSE, echo = 1, fig.align='center', out.width='100%', out.height='500px'}
gradH2O5(TRUE)
knitr::include_graphics("gradH2O5.pdf")
```

## Mean differences of compositional metrics for differentially expressed proteins in osmotic stress (Figure 6)
```{r gradH2O6, message = FALSE, echo = 1, fig.align='center', out.width='100%', out.height='400px'}
gradH2O6(TRUE)
knitr::include_graphics("gradH2O6.pdf")
```

