---
title: "Six plots from salinity gradients paper (2020)"
output: 
  rmarkdown::html_vignette:
    mathjax: null
vignette: >
  %\VignetteIndexEntry{Six plots from salinity gradients paper (2020)}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
library(knitr)
## use pngquant to reduce size of PNG images
knit_hooks$set(pngquant = hook_pngquant)
pngquant <- "--speed=1 --quality=0-25"
# in case pngquant isn't available
if (!nzchar(Sys.which("pngquant"))) pngquant <- NULL 
```

```{r options, include=FALSE}
options(width = 40)
```

```{r HTML, include=FALSE}
## some frequently used HTML expressions
ZC <- "<i>Z</i><sub>C</sub>"
nH2O <- "<i>n</i><sub>H<sub>2</sub>O</sub>"
```

This vignette from the R package [JMDplots](https://github.com/jedick/JMDplots) version `r packageDescription("JMDplots")$Version` contains plots from the paper by [<span title="Distinct trends in chemical composition of proteins from metagenomes in redox and salinity gradients">Dick et al. (2020)</span>](https://doi.org/10.1101/2020.04.09.035022).

```{r JMDplots, message = FALSE, results = "hide"}
library(JMDplots)
```

## Number of reactions for each amino acid in _E. coli_ metabolic reconstruction
Reaction equations used for this calculation are from <a href="https://doi.org/10.1038/msb4100155" title="A genome-scale metabolic reconstruction for Escherichia coli K-12 MG1655 that accounts for 1260 ORFs and thermodynamic information">Feist et al. (2007)</a>.
```{r gradH2O0, message = FALSE, echo = 1}
gradH2O0()
```

## Derivation of stoichiometric hydration state (Figure 1)
```{r gradH2O1, message = FALSE, fig.align='center', out.width='100%', fig.width = 6, fig.height = 6, pngquant = pngquant}
gradH2O1()
```

## `r nH2O`-`r ZC` scatterplots for redox gradients and the Baltic Sea (Figure 2)
```{r gradH2O2, message = FALSE, fig.align='center', out.width='100%', fig.width = 12, fig.height = 5.6, pngquant = pngquant}
gradH2O2()
```

## `r nH2O` for Baltic Sea metagenome and metatranscriptome in different size fractions (Figure 3)
```{r gradH2O3, message = FALSE, fig.align='center', out.width='100%', fig.width = 7, fig.height = 3, pngquant = pngquant}
gradH2O3()
```

## `r nH2O`-`r ZC` for freshwater, marine, and hypersaline environments (Figure 4)
```{r gradH2O4, message = FALSE, fig.align='center', out.width='100%', fig.width = 8, fig.height = 5, pngquant = pngquant}
gradH2O4()
```

## Scatterplots of GRAVY vs ZC and nH2O (Figure 5)
```{r gradH2O5, message = FALSE, fig.align='center', out.width='100%', fig.width = 7, fig.height = 3.2, pngquant = pngquant}
gradH2O5()
```

## `r nH2O`-`r ZC` and GRAVY-pI plots for Baltic Sea and Rodriguez-Brito et al. data (Figure 6)
```{r gradH2O6, message = FALSE, fig.align='center', out.width='100%', fig.width = 8, fig.height = 8, pngquant = pngquant, results = "hide"}
gradH2O6()
```

## Mean differences of compositional metrics for differentially expressed proteins in osmotic stress (Figure 7)
```{r gradH2O7, message = FALSE, fig.align='center', out.width='100%', fig.width = 8, fig.height = 5.2, pngquant = pngquant, results = "hide"}
gradH2O7()
```

