---
title: "Chemical variation of the human microbiome (2023)"
output: 
  rmarkdown::html_vignette:
    mathjax: null
vignette: >
  %\VignetteIndexEntry{Chemical variation of the human microbiome (2023)}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
bibliography: [orp16S.bib, JMDplots.bib]
csl: elementa.csl
---

```{r setup, include=FALSE}
library(knitr)
## use pngquant to reduce size of PNG images
knit_hooks$set(pngquant = hook_pngquant)
pngquant <- "--speed=1 --quality=0-25"
# in case pngquant isn't available
if (!nzchar(Sys.which("pngquant"))) pngquant <- NULL 
```

```{r options, include=FALSE}
options(width = 80)
```

```{r HTML, include=FALSE}
ZC <- "<i>Z</i><sub>C</sub>"
```

This vignette runs the code to make the plots from the following paper:

> Dick JM. 2023. Two dimensions of chemical variation of the human microbiome across body sites and in COVID-19 patients. Manuscript in preparation.
> 
> Preprint available at bioRxiv. doi: [10.1101/2023.02.12.528246](https://doi.org/10.1101/2023.02.12.528246)

This vignette was compiled on `r Sys.Date()` with
**[JMDplots](https://github.com/jedick/JMDplots)** `r packageDescription("JMDplots")$Version`,
**[chem16S](https://github.com/jedick/chem16S)** `r packageDescription("chem16S")$Version`, and
**[canprot](https://github.com/jedick/canprot)** `r packageDescription("canprot")$Version`.

```{r library, message = FALSE, results = "hide"}
library(JMDplots)
```

## Chemical metrics of reference proteomes as a function of oxygen tolerance and comparison with metaproteomes (Figure 1)
```{r sars16S_1, message = FALSE, out.width='100%', fig.align='center', fig.width = 7, fig.height = 5, pngquant = pngquant, dpi = 100}
sars16S_1()
```

**Data sources:** **(A)** *List of Prokaryotes according to their Aerotolerant or Obligate Anaerobic Metabolism* [@MR18]; **(B)** Gut Starch Diet [@MLL+17], Ulcerative Colitis Gut [@TWC+22], Manus Basin Inactive Chimney [@MPB+19], Manus Basin Active Chimneys [@PMM+18;@RYP+14], Soda Lake Biomats [@KTS+17], Mock Communities [@KTS+17], Saanich Inlet [@HTZ+17].

## Chemical variation of microbial proteins across body sites, after viral inactivation treatment, and multi-omics comparison (Figure 2)
```{r sars16S_2, message = FALSE, out.width='100%', fig.width = 8, fig.height = 6, pngquant = pngquant, dpi = 100}
sars16S_2()
```

**Data sources:** **(A)** @BPB+21; **(B)** @BPB+21; **(C)** See Table 1 in manuscript; **(D)** @LLZ+21, @CZH+22, @ZZL+20; **(E)** @TWC+22, @MLL+17, @JZW+22, @GNT+21.

## Differences of chemical metrics of nasopharyngeal, oropharyngeal, and gut microbiomes between controls and COVID-19 patients (Figure 3)
```{r sars16S_3, message = FALSE, out.width='100%', fig.width = 15, fig.height = 12, pngquant = pngquant, dpi = 100}
sars16S_3()
```

**Data sources:** **(A)** See Table 1 in manuscript; **(B)** @HZX+21; **(C)** MAGs [@KWL22], based on data from @ZZL+20 and @YZL+21.

## Overview of chemical variation of the human microbiome inferred from multi-omics datasets (Figure 4)
```{r sars16S_4, message = FALSE, out.width='100%', fig.width = 8, fig.height = 6, pngquant = pngquant, dpi = 100}
sars16S_4()
```

## References
