---
title: "Using thermodynamics to extract geochemical information from genomes (2022)"
output: 
  rmarkdown::html_vignette:
    mathjax: null
vignette: >
  %\VignetteIndexEntry{Using thermodynamics to extract geochemical information from genomes (2022)}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
bibliography: JMDplots.bib
csl: elementa.csl
---

<style>
/* https://gomakethings.com/how-to-break-an-image-out-of-its-parent-container-with-css/ */
@media (min-width: 700px) {
  .full-width {
    left: 50%;
    margin-left: -50vw;
    margin-right: -50vw;
    max-width: 100vw;
    position: relative;
    right: 50%;
    width: 100vw;
  }
}
@media (min-width: 900px) {
  .full-width {
    left: 50vw; /* fallback if needed */
    left: calc(50vw - 100px);
    width: 900px;
    position: relative;
  }
}
</style>

```{r setup, include=FALSE}
library(knitr)
# Use pngquant to reduce size of PNG images
knit_hooks$set(pngquant = hook_pngquant)
pngquant <- "--speed=1 --quality=0-25"
# In case pngquant isn't available
if (!nzchar(Sys.which("pngquant"))) pngquant <- NULL 
```

```{r options, include=FALSE}
options(width = 80)
```

```{r HTML, include=FALSE}
ZC <- "<i>Z</i><sub>C</sub>"
logaH2 <- "log<i>a</i><sub>H<sub>2</sub></sub>"
logfH2 <- "log<i>f</i><sub>H<sub>2</sub></sub>"
logfO2 <- "log<i>f</i><sub>O<sub>2</sub></sub>"
```

This vignette runs the code to make the plots from the following paper:

> Dick JM, Boyer GM, Canovas PA III, Shock EL. 2022. Using thermodynamics to extract geochemical information from genomes. Submitted manuscript.

This vignette was compiled on `r Sys.Date()` with **[JMDplots](https://github.com/jedick/JMDplots)** `r packageDescription("JMDplots")$Version`.

```{r library, message = FALSE}
library(JMDplots)
```

## Chemical analysis of reference proteomes of methanogens reveals adaptation to redox conditions (Figure 1)
```{r utegig1, message = FALSE, out.width='100%', fig.width = 12, fig.height = 8, out.extra = 'class="full-width"', pngquant = pngquant}
utegig1()
```

**Data sources:** Hierarchical clustering tree of methanogenesis genes [@LL18]; `r ZC` calculated for reference proteomes [@DT22] based on RefSeq protein sequences [@OWB+16]. See Table 1 in the paper for references for optimal growth temperature (*T*~opt~).

## Relative stabilities of organic compounds depend on redox conditions (Figure 2)
```{r utegig2, message = FALSE, out.width='100%', fig.width = 7, fig.height = 6, pngquant = pngquant, dpi = 100}
utegig2()
```

**Data sources:** List of organic compounds from @SC10 and TCA cycle metabolites from @CS16. See Table 2 in the paper for references for thermodynamic data.

## Thermodynamic model for methanogen niche partitioning (Figure 3)
```{r utegig3, message = FALSE, out.width='100%', fig.width = 7, fig.height = 5, out.extra = 'class="full-width"', pngquant = "--speed=1 --quality=25-50", dpi = 100}
utegig3()
```

**Data sources:** Lost City *T* and H~2~ concentration [@KKF+05], Rainbow vent fluid mixing model [@SC10], methanogenic and non-methanogenic sediments [@LG88].

## Chemical and thermodynamic analysis of evolutionary divergence along redox gradients (Figure 4 and Table S5)
```{r utegig4, message = FALSE, out.width='100%', fig.width = 9, fig.height = 5, out.extra = 'class="full-width"', dpi = 100}
Ptab <- utegig4()
```

## Parametric and non-parametric tests of comparisons in Fig. 4 (Table S5)
- **p_Tukey**: One-way ANOVA followed by Tukey's Honest Significant Difference test.
- **p_Dunn**: Kruskal-Wallis test followed by Dunn's test with Bonferroni correction for multiple comparisons.
```{r Ptab, message = FALSE, out.width='100%', fig.width = 9, fig.height = 5, out.extra = 'class="full-width"', dpi = 100}
kable(Ptab, row.names = FALSE, digits = 99, format.args = list(scientific = TRUE))
```

**Data sources:** Methanogens [@LL18], Nif-bearing organisms [@PCF+18], Thaumarchaeota [@RFH+19]. Amino acid compositions are from RefSeq proteomes (methanogens and Nif-bearing organisms) or NCBI and IMG genomes (Thaumarchaeota).

## Comparison of `r ZC` of proteomes predicted by Glimmer and downloaded from NCBI (Figure S1)
```{r utegigS1, out.width='60%', fig.align='center', fig.width = 4, fig.height = 4, pngquant = pngquant}
utegigS1()
```

## Association between redox gradients and `r ZC` of proteins and lipids in alkaline Yellowstone hot springs (Figure S2)
```{r utegigS2, out.width='80%', fig.align='center', fig.width = 8.5, fig.height = 4, pngquant = pngquant}
utegigS2()
```

**Data sources:** Proteins [@DS13] and lipids [@BSS+20].

## Relative stabilities of organic compounds at lower activity (Figure S3)
```{r utegigS3, message = FALSE, out.width='100%', fig.align='center', fig.width = 7, fig.height = 3.5, pngquant = pngquant}
utegig2(logact = -6)
```

## `r logaH2`-*T* plots for different organic compound classes (Figure S4)
```{r utegigS4, out.width='80%', fig.align='center', fig.width = 6, fig.height = 8, pngquant = pngquant}
utegigS4()
```

## Convert `r logaH2` to `r logfH2`, `r logfO2`, and Eh
```{r utegig_H2_Eh}
logaH2_15 <- -15
pH <- 7
logK1 <- suppressMessages(subcrt(c("H2", "hydrogen"), c(-1, 1), T = 25)$out$logK)
logfH2 <- logK1 + logaH2_15
logK3 <- suppressMessages(subcrt(c("H2O", "H2", "oxygen"), c(-1, 1, 0.5), T = 25)$out$logK)
logfO2 <- 2 * (logK3 - logaH2_15)
logK3 <- suppressMessages(subcrt(c("H2", "H+", "e-"), c(-1, 2, 2), T = 25)$out$logK)
pe <- (- logK3 - 2 * pH - logaH2_15) / 2
Eh <- convert(pe, "Eh")
print(paste0("Given logaH2 = ", logaH2_15, ", logfH2 = ", round(logfH2, 1), " at equilibrium"))
print(paste0("Given logaH2 = ", logaH2_15, ", logfO2 = ", round(logfO2, 1), " at equilibrium"))
print(paste0("Given logaH2 = ", logaH2_15, " and pH = ", pH, ", Eh = ", round(1000 * Eh), " mV at equilibrium"))
```

## Convert hydrogen partial pressure to `r logaH2`
```{r utegig_H2_H2}
logK <- suppressMessages(subcrt(c("hydrogen", "H2"), c(-1, 1), T = 25)$out$logK)
# For 10^-2 pH2
logpH2 <- -2
print(paste0("Given logpH2 = ", logpH2, ", logaH2 = ", round(logpH2 + logK, 2), " at equilibrium"))
# For 0.55 ppmv H2
logpH2 <- log10(0.55e-6)
print(paste0("Given logpH2 = ", round(logpH2, 2), ", logaH2 = ", round(logpH2 + logK, 2), " at equilibrium"))
```

## References
